---
title: "All Waves EDA"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{All Waves EDA}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## First look 

Initially, we are going to load the data from the package to be using for the
exploratory data analysis.

```{r import}
data("all_waves", package = "BenzoCovid")
```

First, let's take a glimpse and skim the dataset.

```{r explore}
dplyr::glimpse(all_waves)

skimr::skim(all_waves)
```

As we can see in the `skim`, we don't have any variable with missing values.
That's nice!

## Anxiety and depression levels by wave

Now let's create some plots to see the trajectories of the anxiety and depression measures in the sample.

```{r anxiety-and-depression}
all_waves |> 
    BenzoCovid::plot_three_waves("phq_total") +
    ggplot2::labs(x = "Study wave", y = "PHQ-9 total score mean")

all_waves |> 
    BenzoCovid::plot_three_waves("gad_total") +
    ggplot2::labs(x = "Study wave", y = "GAD-7 total score mean")
```

Apparently, we were able to see a decline in both measures, which indicates that people were showing remission of symptoms over time.
