---
title: "Benzodiazepine Use Clustering"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Benzodiazepine Use Clustering}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## The idea

Considering that we have data on the frequency of benzodiazepine use for the three waves in these subjects (*n* = 781), we can try to create clusters referring to the pattern of use in these three moments through an unsupervised machine learning algorithm.

```{r import}
data("all_waves", package = "BenzoCovid")
```

Clustering is a technique in machine learning that attempts to find clusters of
observations within a dataset.
The goal is to find clusters such that the observations within each cluster are
similar to each other, while observations in different clusters are different
from each other. So, we are not predicting a value based on a response/target
variable, but instead, we are grouping the observations.

## Feature transformation

The first step in this process is the transformation of the categorical
variables of the frequency of use of benzodiazepines to numerical variables.
This procedure is necessary because the k-means algorithm only accepts numerical
features as input data.

We also took this step to select only the variables referring to the use of
benzodiazepines in the subset of data.

```{r transform-to-numeric}
benzo_frequencies <- all_waves |> 
    dplyr::mutate(dplyr::across(dplyr::matches("^benzofreq_w[1-3]$"), as.numeric)) |> 
    dplyr::select(dplyr::matches("^benzofreq_w[1-3]$"))
```

## 3D scatterplot

Below is a 3D scatterplot showing the association between benzodiazepine use in
the three assessment waves. In general, there is no great variation in the
use frequency pattern in the sample.

```{r}
fig <- plotly::plot_ly(benzo_frequencies,
                       x = ~benzofreq_w1,
                       y = ~benzofreq_w2,
                       z = ~benzofreq_w3,
                       text = ~paste("Wave 1:", benzofreq_w1,
                                     "<br>Wave 2:", benzofreq_w2,
                                     "<br>Wave 3:", benzofreq_w3),
                       hoverinfo = "text")

fig <- fig |>  plotly::add_markers()

fig <- fig |> 
    plotly::layout(scene = list(
        xaxis = list(title = "Wave 1"),
        yaxis = list(title = "Wave 2"),
        zaxis = list(title = "Wave 3")))


fig
```

```{r}
library(plotly)

data <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/gapminderDataFiveYear.csv")

data_2007 <- data[which(data$year == 2007),]
data_2007 <- data_2007[order(data_2007$continent, data_2007$country),]
data_2007$size <- data_2007$pop
colors <- c('#4AC6B7', '#1972A4', '#965F8A', '#FF7070', '#C61951')

fig <- plot_ly(data_2007, x = ~gdpPercap, y = ~lifeExp, z = ~pop, color = ~continent, size = ~size, colors = colors,
             marker = list(symbol = 'circle', sizemode = 'diameter'), sizes = c(5, 150),
             text = ~paste('Country:', country, '<br>Life Expectancy:', lifeExp, '<br>GDP:', gdpPercap,
                           '<br>Pop.:', pop))
fig <- fig %>% layout(title = 'Life Expectancy v. Per Capita GDP, 2007',
         scene = list(xaxis = list(title = 'GDP per capita (2000 dollars)',
                      gridcolor = 'rgb(255, 255, 255)',
                      range = c(2.003297660701705, 5.191505530708712),
                      type = 'log',
                      zerolinewidth = 1,
                      ticklen = 5,
                      gridwidth = 2),
               yaxis = list(title = 'Life Expectancy (years)',
                      gridcolor = 'rgb(255, 255, 255)',
                      range = c(36.12621671352166, 91.72921793264332),
                      zerolinewidth = 1,
                      ticklen = 5,
                      gridwith = 2),
               zaxis = list(title = 'Population',
                            gridcolor = 'rgb(255, 255, 255)',
                            type = 'log',
                            zerolinewidth = 1,
                            ticklen = 5,
                            gridwith = 2)),
         paper_bgcolor = 'rgb(243, 243, 243)',
         plot_bgcolor = 'rgb(243, 243, 243)')

fig

```


